Geodesic Flow on the Necker Cube Surface

### 2017-06-20 paper.tex
Comments.

The purpose of the following paper is to study the geodesic flow on the periodic surface depicted above, which is formed from an infinite collection of unit squares. We call this surface the Necker cube surface after Louis Albert Necker's "Necker Cube$_{1}$," which became a popular optical illusion. M.C. Escher's adaptation of the cube in Metamorphosis I$_{2}$ briefly features the Necker cube surface (See Figure $\ref{fig:Escher}$).The term Necker Cube is used interchangeably to refer to Escher's surface adaptation and Necker's optical illusion. For the purposes of the following paper, we will be using either Necker Cube or $\mathbf{S}$ to refer to the periodic cube tiling that appears in Escher's work.

@@

The initial (rational) trajectories of any given flow fall into only one of two categories.

@@

\begin{thm*}{Denote $\phi^{\theta}_{t}$ to be the flow in a vector direction $\theta$ to time t on the Necker cube. Then the following is true:}
\begin{enumerate}[label=(\roman*)]
\item If $\theta$ belongs to $\mathfrak{O}$, then $\phi^{\theta}_{t}$ is periodic on $\mathbf{S}$.

@@

\item If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$.

\end{enumerate}
\end{thm*}

@@

\section{Construction the Surface \& Isometric Variant}
\compav{Started up here again.}

@@

\subsection{The Necker Cube Surface and its Flattened Form}
To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square. The three squares have one vertex in common.

@@

\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\chi_{m,n,p} = ([m, m+1]\times[n,n+1]\times\{p\}) 
\\\cup  (\{m+1\}\times[n,n+1]\times[p,p-1]) 
\\\cup ([m,m+1]\times\{n+1\}\times[p,p-1])
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}
\vspace{0.2in}

is the family of these instances based at some integer triple. The surface will be comprised of the members that satisfy the property, $m+n+p=0$. From here on we will refer to the Necker cube surface's embedded form in $\mathbb{R}^{3}$ lying along the plane $z=-(x+y) $ as $\mathbf{S}$. That is to say that

$$\mathbf{S} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}}.$$

$\mathbf{S}$ now resembles the structure in Figure \ref{fig:cubexyz}. Before describing the transformation that takes $\mathbf{S}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges. We will call this surface $\mathbf{U}'$. However, the flattening of the surface takes $\mathbf{S}$ to a translation of $\mathbf{U}'$, we call $\mathbf{U}$ (See $\ref{fig:uprimeu}$). We will denote the metric preserving map from $\mathbf{S}$ to $\mathbf{U}$, by $\Psi$. Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$ due to the nature of the construction of $\mathbf{S}$.

\begin{figure}[H]
\begin{center}
\includesvg[width=1.5 in]{u}\hspace{0.6in}\includesvg[width=1.5in]{uprime}
\end{center}
\caption{$\mathbf{U}$ (left) and $\mathbf{U}'$ (right) differ only by translation.}
\label{fig:uprimeu}
\end{figure}

Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, 
$$\mathbf{P} = \mathbb{R}^{2}-\{(u,v):2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \},$$ with open unit squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin. We will describe this surface as a cross product of the auxiliary plane, $z=0$ and the image of the natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:

@@

Now we will define $\Psi$ piece-wise.

@@

\newpage
\section{The Four-Fold Cover and Translation Surface}
\compav{Started here again.}
The following section is concerned with defining a quotient space on $\mathbf{U}'$ (See $\ref{fig:quotient}$), and using that space to construct our translation surface. In addition we will describe a cover made from four copies of $\mathbf{U}'$ that aligns geodesic flow and behaves as a cover of our translation surface.

@@

The quotient space is a genus 1 surface that resembles a $2\times2$ torus with a unit square removed from it. Near the outer vertex it behaves as a torus might. The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$. None of these vertices are integral multiples of $2\pi$, so in order to create a translation surface we have to take a minimal number of copies of the quotient space so every vertex has a cone angle that is a multiple of $2\pi$.

\newpage
\subsection{Translation Surface}
With $\mathbf{M}$ defined, we can define a branch cover of $\mathbf{M}$ composed of four copies. To each copy we associate a direction of the flow under indices \{0,1,2,3\}. Let $\mathbf{M}_{0} $ denote the quotient space with its vertices removed:

>>>>

The purpose of the following paper is to study the geodesic flow on the periodic surface depicted above, which is formed from an infinite collection of unit squares. We call this surface the Necker cube surface after Louis Albert Necker's ``Necker Cube$_{1}$,'' which became a popular optical illusion. M.C. Escher's adaptation of the cube in Metamorphosis I$_{2}$ briefly features the Necker cube surface (See Figure $\ref{fig:Escher}$).The term Necker Cube is used interchangeably to refer to Escher's surface adaptation and Necker's optical illusion. For the purposes of the following paper, we will be using either Necker Cube or $\mathbf{S}$ to refer to the periodic cube tiling that appears in Escher's work.

@@

The initial (rational) trajectories of any given flow fall into only one of two categories.
\compat{The definitions below seem redundant. Perhaps you want ${\mathcal O}$ and ${\mathcal E}$}

@@

\begin{thm*}{Denote $\phi^{\theta}_{t}$ to be the flow in a vector direction $\theta$ to time t on the Necker cube.
Then the following is true:} \compat{There is no flow in a fixed direction because when trajectories move to new squares they may be traveling different directions.}
\begin{enumerate}[label=(\roman*)]
\item If $\theta$ belongs to $\mathfrak{O}$, then $\phi^{\theta}_{t}$ is periodic on $\mathbf{S}$.

@@

\item If $\theta$ is of the form (x,y) such that x or y are irrational, then $\phi^{\theta}_{t}$ is uniquely ergodic on $\mathbf{S}$. \compat{We should move the discussion of ergodicity outside the theorem.}	
\end{enumerate}
\end{thm*}

@@

\section{Constructing the Surface \& Isometric Variant}
\compav{Started up here again.}

@@

\subsection{The Necker Cube Surface}

To build the surface out of infinitely many squares, we take the union of the instances of three squares, each lying on planes orthogonal to the other two and sharing an edge with exactly one other square. The three squares have one vertex in common.

@@

\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\chi_{m,n,p} = ([m, m+1]\times[n,n+1]\times\{p\}) 
\\\cup  (\{m+1\}\times[n,n+1]\times[p-1,p]) 
\\\cup ([m,m+1]\times\{n+1\}\times[p-1,p])
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}
\vspace{0.2in}
\compat{I would define $A_{m,n,p}$, etc.}

\noindent \compat{Inserted noindent}
is the family of these instances based at some integer triple. The surface will be comprised of the members that satisfy the property, $m+n+p=0$. From here on we will refer to the Necker cube surface's embedded form in $\mathbb{R}^{3}$ lying on the plane $z=-(x+y) $ as $\mathbf{S}$. That is to say that

$$\mathbf{S} =\bigcup\big{\{}\chi_{m,n,p}  : m+n+p=0 \big{\}}.$$
$\mathbf{S}$ now resembles the structure in Figure \ref{fig:cubexyz}. \compat{I would be more definitive. Say $S$ {\em is} the structure...}

\subsection{The Flattened Form}
\compat{I would use as an introductory paragraph: ``We will show that $\mathbf S$ is isometric to a topological quotient of a subset of the plane.''}
Before describing the transformation that takes $\mathbf{S}$ to its flattened form, it would be best to define the relations on a subset of the plane that preserves the identification of edges. We will call this surface $\mathbf{U}'$. However, the flattening of the surface takes $\mathbf{S}$ to a translation of $\mathbf{U}'$, we call $\mathbf{U}$ (See $\ref{fig:uprimeu}$). We will denote the metric preserving map from $\mathbf{S}$ to $\mathbf{U}$, by $\Psi$. Defining these edge identifying relations over $\mathbf{U}'$ is more convenient for us than defining them over $\mathbf{U}$, but defining $\Psi$ is more convenient when the map is taken to $\mathbf{U}$ due to the nature of the construction of $\mathbf{S}$. \compat{I found this paragraph awkward. I would just go ahead and define the region in the plane.}

\begin{figure}[H]
\begin{center}
\includesvg[width=1.5 in]{u}\hspace{0.6in}\includesvg[width=1.5in]{uprime}
\end{center}
\caption{$\mathbf{U}$ (left) and $\mathbf{U}'$ (right) differ only by translation. \compat{Perhaps just use one.}}
\label{fig:uprimeu}
\end{figure}

Now we will describe exactly how to construct $\mathbf{U}'$. Take the subset of the real plane, 
$$\mathbf{P} = \mathbb{R}^{2}-\bigcup_{m,n \in {\mathbb Z}} \{(u,v):2m-\frac{1}{2}<u<2m+\frac{1}{2}, 2n-\frac{1}{2}<v<2n+\frac{1}{2} \},$$ 
\compat{I would write $(m-\frac{1}{2},m+\frac{1}{2}) \times (n-\frac{1}{2},n+\frac{1}{2})$.}
with open unit squares cut out at every even integer pair and let $\mathbf{U}'$ represent the flattened structure with the proper edge identifications with the first "hole" centered at the origin. We will describe this surface as a cross product of the auxiliary plane, $z=0$ and the image of the natural map from $\mathbf{P}$ to its quotient space with the following relation on $\mathbf{P}$:

@@

Now we will define $\Psi$ piecewise.

@@

\newpage
\section{The Four-Fold Cover and Translation Surface}
\compav{Started here again.}
The following section is concerned with defining a quotient of $\mathbf{U}'$ (see Figure $\ref{fig:quotient}$), and using that space to construct a translation surface. In addition we will describe a cover made from four copies of $\mathbf{U}'$ that aligns geodesic flow and behaves as a cover of our translation surface. \compat{Revisit to make sure it summarizes the section}

\subsection{The Euclidean cone surface $\mathbf{M}$}
\compat{I would say: Observe that translations of ${\mathbb R}^2$ by pairs even integers preserves the subset ${\mathbf P}$ and respects the equivalence relation defining ${\mathbf U'}$. Thus this gives an action of $(2{\mathbb Z})^2$ on ${\mathbf U}'$. We define ${\mathbf M}$ to be the quotient. It is depicted in Figure ...}

@@

The quotient space is a genus 1 surface that resembles a $2\times2$ torus with a unit square removed from it. Near the outer vertex it behaves as a torus might. The inner vertices $c_{1},c_{3}$ have cone angle $\frac{3\pi}{2}$, and $c_{2}$ has a cone angle of $3\pi$. None of these vertices are integral multiples of $2\pi$, so in order to create a translation surface we have to take a minimal number of copies of the quotient space so every vertex has a cone angle that is a multiple of $2\pi$. \compat{Move the discussion of building the translation surface to the next section. Otherwise this paragraph is good.}

\newpage
\subsection{Translation Surface}
\compat{You need to build the surface out of pieces rather than out of ``copies of $\mathbf{M}$.''} 
We will now define a translation which is a four-fold branch cover of $\mathbf{M}$. To each copy we associate a direction of the flow under indices \{0,1,2,3\}. Let $\mathbf{M}_{0} $ denote the quotient space with its vertices removed:
